---
title: "Sex offense, revise"
author: "Rongbin Xu"
date: "2020/5/14"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
This is a simple tutorial of time-stratified case-crossover design based on the data described in
Xu, R., et al. (2021). "Association between ambient temperature and sex offense: A case-crossover study in seven large US cities, 2007–2017." Sustainable Cities and Society 69: 102828.


## load parckages
```{r}
rm(list = ls())
library(tidyverse)## for data processing
library(dlnm)## for lag models
library(survival)## for conditional logistic regression
library(splines) ## for non-linear splines
library(mvmeta) ## for meta-regression
library(lubridate) ## processing dates
```



## Step 1: process data into case-crossover format
```{r}
## load data with each crime records
load("Sample_data.Rdata")


## run a funtion for tranforming record data to case-crossover data
f_case<-function(mydata){
mydata$patient<-1
mydata$id<-1:nrow(mydata)

##find all possible candidate
d<-c(-21,-14,-7,7,14,21)
mydata2<-mydata
for (i in 1:6) {
  da<-mydata%>%
    mutate(date=date+d[i])%>%
    mutate(patient=0)
  mydata2<-bind_rows(mydata2,da)
  rm(da)
}

## filter only controls within one month
case_month<-mydata2%>%
  mutate(m=month(date))%>%
  filter(patient==1)%>%
  group_by(id)%>%
  summarise(m_case=m)

mydata2<-mydata2%>%
  mutate(m=month(date))%>%
  left_join(case_month)%>%
  filter(m==m_case)%>%
  select(-m,-m_case)

## return results
mydata2
}


## Try an example using the first two rows
f_case(crime_data[1,])%>%View()
f_case(crime_data[123,])## try a random row


## Make the data for case-crossover design
my_crossover<-f_case(crime_data)
```

## Process climate data
```{r}
climate<-read.csv("USA_climate_data.csv")
climate<-climate[,-1]

###temperature lag
for (i in 1:30) { ## for short-term effects, up to 30 days lag is enough
  climate<-climate%>%
    group_by(city)%>%
    mutate(tmeanlag=lag(tmean,i))%>% ## tmean is the tmean_lag0, the current day temperature
    ungroup()
  
  names(climate)[ncol(climate)]<-paste("tmean_lag",i,sep = "")
  print(i)
}


###precipitation lag
for (i in 1:30) {
  climate<-climate%>%
    group_by(city)%>%
    mutate(preciplag=lag(precipitation,i))%>%
    ungroup()
  
  names(climate)[ncol(climate)]<-paste("precipitation_lag",i,sep = "")
  print(i)
}


###humidity lag
for (i in 1:30) {
  climate<-climate%>%
    group_by(city)%>%
    mutate(humidity.lag=lag(humidity,i))%>%
    ungroup()
  
  names(climate)[ncol(climate)]<-paste("humidity_lag",i,sep = "")
  print(i)
}


## an example of creating complete time-series data
complete_time.series<-expand.grid(city=c("A","B","C"),
            date=seq.Date(from = as.Date("2007-01-01"),
                          to=as.Date("2012-12-31"),by=1)
            )%>%
  arrange(city,date)

```
## Link case-crossover data with climate data
```{r}
str(climate)
str(my_crossover)

climate$date<-as.Date(climate$date)
mydata<-my_crossover%>%
  left_join(climate,by=c("city","date"))
```


## Case-crossover design based on dlnm
```{r}

## build a cross-basis for temperature
names(mydata)
tem.basis <- crossbasis(mydata[,c(7,15:22)],argvar=list(fun="lin"),
                          arglag=list(fun="ns",knots=logknots(8,nk=1)),lag=8)

## note  mydata[,c(7,15:22)] we chose the columns of tmean lag0 to lag8


## basis for humidity and precipitation
h.basis<-crossbasis(mydata[,c(10,75:82)],argvar=list(fun="ns",df=3),
                      arglag=list(fun="ns",knots=logknots(8,nk=1)),lag=8)

p.basis<-crossbasis(mydata[,c(6,45:52)],argvar=list(fun="ns",df=3),
                      arglag=list(fun="ns",knots=logknots(8,nk=1)),lag=8)

p.basis1<-onebasis(mydata$precipitation,fun = "ns",df=3)## an example of adjusting a variable without lag using non-linear model


## run conditional logistic model
model<-clogit(patient~tem.basis+h.basis+p.basis+holiday+strata(id),method = "exact",data = mydata)## method="breslow"




BIC(model)## extract BIC value, a value evaluating the model performace, the lower the bettter, this would be useful for selecting df

## present model
tem.pred <- crosspred(tem.basis,model,at=-200:320/10,by=0.1,cen=-20,bylag = 0.1) 
### cen is the reference temperture, can be self-defined
### at=-20:32 is a self-defined temperature range, here I used -200:320/10, so it can allow temperature at 0.1 degree units

plot(tem.pred,xlab="Temperature",ylab="Lag (days)",zlab="Odds ratio")## 3D plot
plot(tem.pred,"overall",ylab="Odds ratio",xlab="Temperature",col="red")## overall cumulative associations for all lags
plot(tem.pred,var = 15,exp=T,ylab="Odds ratio",xlab="Lag (days)",col="red")## lag pattern for 15 vs 10
plot(tem.pred,lag = 1,exp=T,ylab="Odds ratio",xlab="Temperature",col="red")## effect for lag=1
```
## Extract effect estimates
```{r}
 ## try use "tem.pred$" this to explor the structure of tem.pred

tem.pred$allRRfit## cumulative RR or OR, the RR at 10(the cen value defined above) is 1, other RRs are RRs relative to this temperature
tem.pred$allRRlow## low bound of 95%CI for the cumulative RR
tem.pred$allRRhigh ## upbound of the 95%CI for the cumulative RR

tem.pred$allfit## Cumulative Beta, exp(tem.pred$allfit)=tem.pred$allRRfit
exp(tem.pred$allfit)-tem.pred$allRRfit ## should equal to 0
tem.pred$allse ## cumulative stardard error(SE), Beta and SE is useful to get the P-value


## Extract effect estimates for every 5 degree increase, i.e., the estimates at 15 degree relative to 10 degrees
tem.pred$allRRfit[36]
tem.pred$allRRlow[36]
tem.pred$allRRhigh[36]

Beta<-tem.pred$allfit[36]  
Se<-tem.pred$allse[36]  

## Beta and Se can be used to calculate P-values
P.value<-(1-pnorm(abs(Beta/Se),0,1))*2 

```
## build a function for case-crossover analyses
```{r}
f_model<-function(da){
  tem.basis <- crossbasis(da[,c(7,15:22)],argvar=list(fun="lin"),
                          arglag=list(fun="ns",knots=logknots(8,nk=1)),lag=8)

## note  mydata[,c(7,15:22)] we chose the columns of tmean lag0 to lag8


## basis for humidity and precipitation
h.basis<-crossbasis(da[,c(10,75:82)],argvar=list(fun="ns",df=3),
                      arglag=list(fun="ns",knots=logknots(8,nk=1)),lag=8)

p.basis<-crossbasis(da[,c(6,45:52)],argvar=list(fun="ns",df=3),
                      arglag=list(fun="ns",knots=logknots(8,nk=1)),lag=8)


## run conditional logistic model
model<-clogit(patient~tem.basis+h.basis+p.basis+holiday+strata(id),method = "exact",data = da)## method="breslow"

## present model
tem.pred <- crosspred(tem.basis,model,at=-20:32,by=0.1,cen=10,bylag = 0.1) 
tem.pred
}

## a loop for extracting effect estimates
result<-data.frame(city=unique(mydata$city),Beta=NA,Se=NA,P.value=NA,RR=NA,RR.low=NA,RR.up=NA)

for (i in 1:7) {
  ## extract data for a specific city
  data<-mydata%>%filter(city==result$city[i])
  
  ## run model for this city
  m<-f_model(data)
  
  ## save results
  result[i,2]<-m$allfit[36]
  result[i,3]<-m$allse[36]
  result[i,5]<-m$allRRfit[36]
  result[i,6]<-m$allRRlow[36]
  result[i,7]<-m$allRRhigh[36]
  
  ## monitoring progress
  print(i)

}

## calculate P-value
result<-result%>%
  mutate(P.value=(1-pnorm(abs(Beta/Se),0,1))*2 )

```

## Meta regression
```{r}
str(result$city)
result$city<-as.factor(result$city)
levels(result$city)
result$city<-factor(result$city,levels = result$city)

meta_model<-mvmeta(Beta~city,S=Se^2,data = result,method = "ml")## also can try method="reml"
summary(meta_model)
summary(meta_model)$coefficients


## save P-value for difference
result$P.diff[2:7]<-summary(meta_model)$coefficients[2:7,4]

## save results
write.csv(result,"city_specific_estimates.csv")
```

